---
title: "snRNA-seq HER2 Analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    fig_width: 10
    fig_height: 8
---

1. Restore packages
```{r setup, include=False}
set.seed(42)

# 1. Install and load required R packages
required_packages <- c("Seurat", "tidyverse", "here", "sceasy", "reticulate")

for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
}

# 2. Setup Python environment with anndata
if (!reticulate::py_module_available("anndata")) {
  message("Installing anndata Python package...")
  reticulate::py_install("anndata", pip = TRUE)
  message("anndata installation complete!")
}

# 3. Verify Python setup
stopifnot("anndata not available" = reticulate::py_module_available("anndata"))

# 4. Create directories
paths <- list(
  raw = here("data", "raw"),
  results = here("results")
)

walk(paths, ~dir.create(.x, showWarnings = FALSE, recursive = TRUE))
```

2. Load knitr options
```{r}
knitr::opts_chunk$set(
  echo = TRUE,           # Show code
  message = FALSE,       # Hide package load messages
  warning = FALSE,       # Hide warnings
  error = FALSE,         # Stop on error (don't keep knitting)
  fig.align = "center",
  fig.width = 8,         # Default figure width (inches)
  fig.height = 6,        # Default figure height
  dpi = 300,             # Publication-quality resolution
  cache = TRUE,          # Cache expensive chunks
  cache.lazy = FALSE     # Safer for large objects (Seurat, etc.)
)
```

GSE169379_MIBC_snSeq.h5ad: 25 treatment-naive muscle-invasive bladder cancer (MIBC) patients
- Contains CDH12+ epithelial subpopulation that predicts poor outcome after surgery/chemo but good response to immunotherapy

GSE169379_non_tumor_snSeq.h5ad: 3 histologically-normal bladder tissue samples


```{r}

# muscle-invasive bladder cancer
mibc_h5ad_path <- here("data", "GSE169379_MIBC_snSeq.h5ad")

#normal bladder
normal_h5ad_path <- here("data", "GSE169379_non_tumor_snSeq.h5ad")

library(sceasy)

mibc   <- convertFormat(mibc_h5ad_path, from = "anndata", to = "seurat")
normal <- convertFormat(normal_h5ad_path, from = "anndata", to = "seurat")

#optional subset to make processing faster (temporarily)
mibc <- subset(mibc, cells=sample(Cells(mibc), size=floor(0.01*ncol(mibc))))

```

```{r}
mibc_epithelial <- subset(mibc, subset=celltype=="Epithelial")
normal_known <- subset(normal, subset= SubType_Normal != "Normal_Unk") #filter out unknown cell types

```


Inputs for inferCNV
```{r}

```



```{r}

infercnv_obj = CreateInfercnvObject(raw_counts_matrix="singleCell.counts.matrix",
                                    annotations_file="cellAnnotations.txt",
                                    delim="\t",
                                    gene_order_file="gene_ordering_file.txt",
                                    ref_group_names=c("normal"))

# perform infercnv operations to reveal cnv signal
infercnv_obj = infercnv::run(infercnv_obj,
                             cutoff=1,  # use 1 for smart-seq, 0.1 for 10x-genomics
                             out_dir="output_dir",  # dir is auto-created for storing outputs
                             cluster_by_groups=T,   # cluster
                             denoise=T,
                             HMM=T
                             )

```


Module Scoring
```{r}

# genes associated w/ HER2-positive status
her2_positive <- c("MEAF6", "HMGCS1", "ZFYVE19", "GIPC2", "FGG", "ZWILCH", "TUT1", "POLE", "ZPR1", "CDKL2", "RIPOR1", "NUP58", "PPIL1", "PSMA4", "XRN1", "HIST2H2AC", "UGT1A4", "SRP68", "SNX21", "ABCE1", "ERBB2", "CCT6B", "TLR7", "GRHL2")
# ZNF259 -> ZPR1
# FAM65A -> RIPOR1
# NUPL1 -> NUP58

# genes associated w/ HER2-negative status
her2_negative <-c("WDR3", "EIF3G", "UQCRC2", "ZCCHC9", "RRM2", "NDUFAF7", "PRAME", "RPF2", "MED7", "IQCF3", "RPL14", "HEATR1", "RPS20", "CMTM1", "RPS17", "RPS15A")

```

```{r}

mibc_epithelial <- AddModuleScore(
  mibc_epithelial,
  features = list(her2_positive, her2_negative),
  name = c("HER2_pos_score", "HER2_neg_score")
)

mibc_epithelial$HER2_combined <- 
  mibc_epithelial$HER2_pos_score1 - mibc_epithelial$HER2_neg_score2

```

```{r}

scores <- mibc_epithelial$HER2_combined
lo_thresh <- quantile(scores, 0.30)  # 30th percentile value
hi_thresh <- quantile(scores, 0.70)  # 70th percentile value

mibc_epithelial$HER2_bin <- NA # all cells = NA
mibc_epithelial$HER2_bin[scores <= lo_thresh] <- "HER2lo" #bottom30
mibc_epithelial$HER2_bin[scores >= hi_thresh] <- "HER2hi" #top30
# middle 40 stay NA

binned_epithelial <- subset(mibc_epithelial, subset = HER2_bin %in% c("HER2hi", "HER2lo")) #remove middle40 cells; keep top30 and bottom30 only

Idents(binned_epithelial) <- "HER2_bin"

```


DGE between HER2hi and HER2lo
- Trying to find transcriptional differences b/w HER2hi and HER2lo epithelial cells
```{r}

dge_results <- FindMarkers(binned_epithelial, ident.1="HER2hi", ident.2="HER2lo", test.use="wilcox", logfc.threshold=0.25)

```
Filter for significant DGEs
```{r}
library(ggplot2)
library(dplyr)

dge_results <- dge_results %>%
  tibble::rownames_to_column("gene") %>%
  mutate(
    significant = p_val_adj < 0.05 & abs(avg_log2FC) > 0.25,
    direction   = case_when(
      avg_log2FC > 0.25  & p_val_adj < 0.05 ~ "Up in HER2hi",
      avg_log2FC < -0.25 & p_val_adj < 0.05 ~ "Up in HER2lo",
      TRUE ~ "Not significant"
    )
  )

# Save DEGs
write.csv(dge_results, here("results", "DEGs_HER2hi_vs_lo.csv"), row.names = FALSE)

```



Volcano plot visualization
```{r}

table(dge_results$direction)

# --- Volcano plot ---
ggplot(dge_results, aes(x = avg_log2FC, y = -log10(p_val_adj), color = direction)) +
  geom_point(alpha = 0.6, size = 1.5) +
  scale_color_manual(values = c(
    "Up in HER2hi"    = "#e74c3c",
    "Up in HER2lo"    = "#3498db",
    "Not significant" = "grey70"
  )) +
  geom_vline(xintercept = c(-0.25, 0.25), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05),   linetype = "dashed") +
  # Label top 10 genes per direction
  ggrepel::geom_text_repel(
    data = dge_results %>%
      filter(significant) %>%
      group_by(direction) %>%
      slice_max(abs(avg_log2FC), n = 10),
    aes(label = gene), size = 3, max.overlaps = 20
  ) +
  labs(
    title = "HER2hi vs HER2lo Epithelial Cells",
    x     = "Log2 Fold Change",
    y     = "-log10 Adjusted P-value",
    color = NULL
  ) +
  theme_classic()

```
Pathway Enrichment
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)

her2hi_genes <- dge_results %>% filter(direction == "Up in HER2hi") %>% pull(gene)
her2lo_genes <- dge_results %>% filter(direction == "Up in HER2lo") %>% pull(gene)

run_go <- function(genes, label) {
  enrichGO(
    gene          = genes,
    OrgDb         = org.Hs.eg.db,
    keyType       = "SYMBOL",   # our genes are already gene symbols
    ont           = "BP",       # Biological Process
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05,
    qvalueCutoff  = 0.2
  )
}

go_her2hi <- run_go(her2hi_genes, "HER2hi")
go_her2lo <- run_go(her2lo_genes, "HER2lo")

# Dot plots
dotplot(go_her2hi, showCategory = 15, title = "Pathways enriched in HER2hi")
dotplot(go_her2lo, showCategory = 15, title = "Pathways enriched in HER2lo")

# Save
write.csv(as.data.frame(go_her2hi), here("results", "GO_HER2hi.csv"), row.names = FALSE)
write.csv(as.data.frame(go_her2lo), here("results", "GO_HER2lo.csv"), row.names = FALSE)

```
