---
title: "snRNA-seq HER2 Analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    fig_width: 10
    fig_height: 8
---

1. Restore packages
```{r}
if (!require("renv", character.only = TRUE, quietly = TRUE)) {
  install.packages("renv")
}

if (!file.exists("renv.lock")) {
  stop("No renv.lock found. Run renv::snapshot() on source machine.")
}

renv::restore()
```

2. Load knitr options
```{r}
knitr::opts_chunk$set(
  echo = TRUE,           # Show code
  message = FALSE,       # Hide package load messages
  warning = FALSE,       # Hide warnings
  error = FALSE,         # Stop on error (don't keep knitting)
  fig.align = "center",
  fig.width = 8,         # Default figure width (inches)
  fig.height = 6,        # Default figure height
  dpi = 300,             # Publication-quality resolution
  cache = TRUE,          # Cache expensive chunks
  cache.lazy = FALSE     # Safer for large objects (Seurat, etc.)
)
```

3. Load libraries
```{r}
set.seed(42)

library(Seurat)
library(tidyverse)
library(here)

paths <- list(
  raw = here("data", "raw"),
  results = here("results")
)

walk(paths, ~dir.create(.x, showWarnings = FALSE, recursive = TRUE))
```

GSE169379_MIBC_snSeq.h5ad: 25 treatment-naive muscle-invasive bladder cancer (MIBC) patients
- Contains CDH12+ epithelial subpopulation that predicts poor outcome after surgery/chemo but good response to immunotherapy

GSE169379_non_tumor_snSeq.h5ad: 3 histologically-normal bladder tissue samples


```{r}

# muscle-invasive bladder cancer
mibc_h5ad_path <- here("data", "GSE169379_MIBC_snSeq.h5ad")

#normal bladder
normal_h5ad_path <- here("data", "GSE169379_non_tumor_snSeq.h5ad")

library(sceasy)

mibc   <- convertFormat(mibc_h5ad_path, from = "anndata", to = "seurat")
normal <- convertFormat(normal_h5ad_path, from = "anndata", to = "seurat")

```

```{r}
mibc_epithelial <- subset(mibc, subset=celltype=="Epithelial")
normal_known <- subset(normal, subset= SubType_Normal != "Normal_Unk") #filter out unknown cell types

```


Inputs for inferCNV
```{r}

```



```{r}

infercnv_obj = CreateInfercnvObject(raw_counts_matrix="singleCell.counts.matrix",
                                    annotations_file="cellAnnotations.txt",
                                    delim="\t",
                                    gene_order_file="gene_ordering_file.txt",
                                    ref_group_names=c("normal"))

# perform infercnv operations to reveal cnv signal
infercnv_obj = infercnv::run(infercnv_obj,
                             cutoff=1,  # use 1 for smart-seq, 0.1 for 10x-genomics
                             out_dir="output_dir",  # dir is auto-created for storing outputs
                             cluster_by_groups=T,   # cluster
                             denoise=T,
                             HMM=T
                             )

```


Module Scoring
```{r}
her2_positive <- c("MEAF6", "HMGCS1", "ZFYVE19", "GIPC2", "FGG", "ZWILCH", "TUT1", "POLE", "ZPR1", "CDKL2", "RIPOR1", "NUP58", "PPIL1", "PSMA4", "XRN1", "HIST2H2AC", "UGT1A4", "SRP68", "SNX21", "ABCE1", "ERBB2", "CCT6B", "TLR7", "GRHL2")
# ZNF259 -> ZPR1
# FAM65A -> RIPOR1
# NUPL1 -> NUP58


her2_negative <-c("WDR3", "EIF3G", "UQCRC2", "ZCCHC9", "RRM2", "NDUFAF7", "PRAME", "RPF2", "MED7", "IQCF3", "RPL14", "HEATR1", "RPS20", "CMTM1", "RPS17", "RPS15A")

```

```{r}
AddModuleScore(
  object,
  features,
  pool = NULL,
  nbin = 24,
  ctrl = 100,
  k = FALSE,
  assay = NULL,
  name = "Cluster",
  seed = 1,
  search = FALSE,
  slot = "data",
  ...
)


```




