---
title: "snRNA-seq HER2 Analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    fig_width: 10
    fig_height: 8
---

1. Restore packages
```{r setup, include=False}
set.seed(42)

# 1. Install and load required R packages

# Install (and load) CRAN packages
pacman::p_load(Seurat, tidyverse, reticulate, ggplot2, ggrepel)

# Install (and load) bioconductor packages
BiocManager::install(c("clusterProfiler", "enrichplot", "org.Hs.eg.db", "sceasy"))
pacman::p_load(clusterProfiler, enrichplot, org.Hs.eg.db, sceasy)

# Install and load github packages
pacman::p_load_gh("saeyslab/nichenetr")

# Python setup (runs only if anndata is missing)
if (!reticulate::py_module_available("anndata")) reticulate::py_install("anndata")

#5 Setup Python aand verify anndata
library(reticulate)


if (!py_module_available("anndata")) {
  message("anndata not found. Setting up Python environment...")
  
  # Option 1: Install in reticulate's default environment
  tryCatch({
    py_install("anndata")
    message("Successfully installed anndata")
  }, error = function(e) {
    message("Failed to install in default environment: ", e$message)
    message("Trying alternative approach...")
    
    # Option 2: Create a dedicated conda environment
    if (requireNamespace("reticulate", quietly = TRUE)) {
      conda_env <- "r-reticulate"
      if (!conda_env %in% conda_list()$name) {
        conda_create(conda_env, packages = "python=3.9")
      }
      use_condaenv(conda_env, required = TRUE)
      py_install("anndata", envname = conda_env)
    }
  })
}

# Verify anndata is now available
stopifnot("anndata is still not available. Please install manually: pip install anndata" = 
          py_module_available("anndata"))

message("✓ All packages and Python modules ready")

# 6 Create directories
paths <- list(
  raw = here("data", "raw"),
  results = here("results")
)

walk(paths, ~dir.create(.x, showWarnings = FALSE, recursive = TRUE))
```

2. Load knitr options
```{r}
knitr::opts_chunk$set(
  echo = TRUE,           # Show code
  message = FALSE,       # Hide package load messages
  warning = FALSE,       # Hide warnings
  error = FALSE,         # Stop on error (don't keep knitting)
  fig.align = "center",
  # fig.width = 8,         # Default figure width (inches)
  # fig.height = 6,        # Default figure height
  dpi = 300,             # Publication-quality resolution
  cache = TRUE,          # Cache expensive chunks
  cache.lazy = FALSE     # Safer for large objects (Seurat, etc.)
)
```

GSE169379_MIBC_snSeq.h5ad: 25 treatment-naive muscle-invasive bladder cancer (MIBC) patients
- Contains CDH12+ epithelial subpopulation that predicts poor outcome after surgery/chemo but good response to immunotherapy

GSE169379_non_tumor_snSeq.h5ad: 3 histologically-normal bladder tissue samples


```{r}

# muscle-invasive bladder cancer
mibc_h5ad_path <- here("data", "GSE169379_MIBC_snSeq.h5ad")

#normal bladder
normal_h5ad_path <- here("data", "GSE169379_non_tumor_snSeq.h5ad")

library(sceasy)

mibc   <- convertFormat(mibc_h5ad_path, from = "anndata", to = "seurat")
normal <- convertFormat(normal_h5ad_path, from = "anndata", to = "seurat")

#optional subset to make processing faster (temporarily)
mibc <- subset(mibc, cells=sample(Cells(mibc), size=floor(0.01*ncol(mibc))))

```

Visualize UMAP and tsne
```{r}
mibc <- NormalizeData(mibc)
mibc <- FindVariableFeatures(mibc)
mibc <- ScaleData(mibc)
mibc <- RunPCA(mibc)
mibc <- RunUMAP(mibc, dims = 1:30)
mibc <- RunTSNE(mibc, dims = 1:30)

normal <- NormalizeData(normal)
normal <- FindVariableFeatures(normal)
normal <- ScaleData(normal)
normal <- RunPCA(normal)
normal <- RunUMAP(normal, dims = 1:30)
normal <- RunTSNE(normal)
```

```{r}
DimPlot(mibc, reduction = "umap", group.by = "initial_celltype")

```

```{r}
mibc_epithelial <- subset(mibc, subset=celltype=="Epithelial")
normal_known <- subset(normal, subset= SubType_Normal != "Normal_Unk") #filter out unknown cell types

```


Inputs for inferCNV
```{r}

```



```{r}

# infercnv_obj = CreateInfercnvObject(raw_counts_matrix="singleCell.counts.matrix",
#                                     annotations_file="cellAnnotations.txt",
#                                     delim="\t",
#                                     gene_order_file="gene_ordering_file.txt",
#                                     ref_group_names=c("normal"))
# 
# # perform infercnv operations to reveal cnv signal
# infercnv_obj = infercnv::run(infercnv_obj,
#                              cutoff=1,  # use 1 for smart-seq, 0.1 for 10x-genomics
#                              out_dir="output_dir",  # dir is auto-created for storing outputs
#                              cluster_by_groups=T,   # cluster
#                              denoise=T,
#                              HMM=T
#                              )

```


Module Scoring
```{r}

# genes associated w/ HER2-positive status
her2_positive <- c("MEAF6", "HMGCS1", "ZFYVE19", "GIPC2", "FGG", "ZWILCH", "TUT1", "POLE", "ZPR1", "CDKL2", "RIPOR1", "NUP58", "PPIL1", "PSMA4", "XRN1", "HIST2H2AC", "UGT1A4", "SRP68", "SNX21", "ABCE1", "ERBB2", "CCT6B", "TLR7", "GRHL2")
# ZNF259 -> ZPR1
# FAM65A -> RIPOR1
# NUPL1 -> NUP58

# genes associated w/ HER2-negative status
her2_negative <-c("WDR3", "EIF3G", "UQCRC2", "ZCCHC9", "RRM2", "NDUFAF7", "PRAME", "RPF2", "MED7", "IQCF3", "RPL14", "HEATR1", "RPS20", "CMTM1", "RPS17", "RPS15A")

```

```{r}

mibc_epithelial <- AddModuleScore(
  mibc_epithelial,
  features = list(her2_positive, her2_negative),
  name = c("HER2_pos_score", "HER2_neg_score")
)

mibc_epithelial$HER2_combined <- 
  mibc_epithelial$HER2_pos_score1 - mibc_epithelial$HER2_neg_score2

```

```{r}

scores <- mibc_epithelial$HER2_combined
lo_thresh <- quantile(scores, 0.30)  # 30th percentile value
hi_thresh <- quantile(scores, 0.70)  # 70th percentile value

mibc_epithelial$HER2_bin <- NA # all cells = NA
mibc_epithelial$HER2_bin[scores <= lo_thresh] <- "HER2lo" #bottom30
mibc_epithelial$HER2_bin[scores >= hi_thresh] <- "HER2hi" #top30
# middle 40 stay NA

binned_epithelial <- subset(mibc_epithelial, subset = HER2_bin %in% c("HER2hi", "HER2lo")) #remove middle40 cells; keep top30 and bottom30 only

Idents(binned_epithelial) <- "HER2_bin"

```


DGE between HER2hi and HER2lo
- Trying to find transcriptional differences b/w HER2hi and HER2lo epithelial cells
```{r}

dge_results <- FindMarkers(binned_epithelial, ident.1="HER2hi", ident.2="HER2lo", test.use="wilcox", logfc.threshold=0.25)

```
Filter for significant DGEs
```{r}
library(ggplot2)
library(dplyr)

dge_results <- dge_results %>%
  tibble::rownames_to_column("gene") %>%
  mutate(
    significant = p_val_adj < 0.05 & abs(avg_log2FC) > 0.25,
    direction   = case_when(
      avg_log2FC > 0.25  & p_val_adj < 0.05 ~ "Up in HER2hi",
      avg_log2FC < -0.25 & p_val_adj < 0.05 ~ "Up in HER2lo",
      TRUE ~ "Not significant"
    )
  )

# Save DEGs
write.csv(dge_results, here("results", "DEGs_HER2hi_vs_lo.csv"), row.names = FALSE)

```



Volcano plot visualization
```{r}

table(dge_results$direction)

# --- Volcano plot ---
ggplot(dge_results, aes(x = avg_log2FC, y = -log10(p_val_adj), color = direction)) +
  geom_point(alpha = 0.6, size = 1.5) +
  scale_color_manual(values = c(
    "Up in HER2hi"    = "#e74c3c",
    "Up in HER2lo"    = "#3498db",
    "Not significant" = "grey70"
  )) +
  geom_vline(xintercept = c(-0.25, 0.25), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05),   linetype = "dashed") +
  # Label top 10 genes per direction
  ggrepel::geom_text_repel(
    data = dge_results %>%
      filter(significant) %>%
      group_by(direction) %>%
      slice_max(abs(avg_log2FC), n = 10),
    aes(label = gene), size = 3, max.overlaps = 20
  ) +
  labs(
    title = "HER2hi vs HER2lo Epithelial Cells",
    x     = "Log2 Fold Change",
    y     = "-log10 Adjusted P-value",
    color = NULL
  ) +
  theme_classic()

```
Pathway Enrichment
```{r}
library(clusterProfiler)
library(org.Hs.eg.db)

her2hi_genes <- dge_results %>% filter(direction == "Up in HER2hi") %>% pull(gene)
her2lo_genes <- dge_results %>% filter(direction == "Up in HER2lo") %>% pull(gene)

run_go <- function(genes, label) {
  enrichGO(
    gene          = genes,
    OrgDb         = org.Hs.eg.db,
    keyType       = "SYMBOL",   # our genes are already gene symbols
    ont           = "BP",       # Biological Process
    pAdjustMethod = "BH",
    pvalueCutoff  = 0.05,
    qvalueCutoff  = 0.2
  )
}

go_her2hi <- run_go(her2hi_genes, "HER2hi")
go_her2lo <- run_go(her2lo_genes, "HER2lo")
```

```{r}
# Show plot (will have overlapping text)
dotplot(go_her2hi, showCategory = 15, title = "Pathways enriched in HER2hi")
dotplot(go_her2lo, showCategory = 15, title = "Pathways enriched in HER2lo")

# SAVE AS PDF
pdf("HER2_dotplots.pdf", width = 8, height = 14)  # save PDF
  dotplot(go_her2hi, showCategory = 15, title = "Pathways enriched in HER2hi")
  dotplot(go_her2lo, showCategory = 15, title = "Pathways enriched in HER2lo")
dev.off()

```

NICHENET
```{r}
library(nichenetr)

mibc_epithelial@meta.data$initial_celltype %>% table()

#Read in NicheNet's own ligand-receptor networks
lr_network <- readRDS(url("https://zenodo.org/record/7074291/files/lr_network_human_21122021.rds"))
  ligand_target_matrix <- readRDS(url("https://zenodo.org/record/7074291/files/ligand_target_matrix_nsga2r_final.rds"))
  weighted_networks <- readRDS(url("https://zenodo.org/record/7074291/files/weighted_networks_nsga2r_final.rds"))
```

NICHENET AGNOSTIC
```{r}
nichenet_output_agnostic <- nichenet_seuratobj_aggregate(
  seurat_obj = mibc_epithelial, 
  sender = "undefined",
  receiver = "Endothelial", 
  condition_colname = "aggregate",
  condition_oi = "LCMV",
  condition_reference = "SS",
  expression_pct = 0.05,
  ligand_target_matrix = ligand_target_matrix,
  lr_network = lr_network,
  weighted_networks = weighted_networks
  )

```

```{r}
# test

```